Generalized Tractability

### 2023-05-10 GeneralizedTractability_230501.tex
Updates from Overleaf

Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},d',\vp)$ are non-decreasing in $i$ for any $d'\in\naturals$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},d',\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},d',\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},d',\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) } \le \frac{1}{T(\varepsilon^{-1},d',\vp)} \\ \forall d', n, N \in \naturals, \ N \le n.
\end{multline}

Moreover, given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity via \eqref{eq:comp_def}:
\begin{align} %\label{eq:comp_equiv}
	\nonumber
	\comp(\varepsilon, d) & = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
	\nonumber
	& = \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},d',\vp)}\le \frac{1}{T(\varepsilon^{-1},d',\vp) }\biggr\} \qquad  \forall d'\in \naturals\\
	\nonumber
	& \le  \min \biggl\{n \in \natzero : n \ge N - 1 + T(\varepsilon^{-1},d',\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \biggr\} \;  \forall  d', N \in \naturals, \quad \text{by \eqref{eq:implicate_a}}. 
	\intertext{Note that in this last expression, $n \ge N$ is enforced since $n$ is an integer strictly greater than  $N-1$ by definition.  This then implies that}
	\comp(\varepsilon, d) 
	& \le  T(\varepsilon^{-1},d',\vp) \Biggl [\frac{N}{T(\varepsilon^{-1},d',\vp)}  +  \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \Biggr] \quad \forall  d', N \in \naturals. \label{eq:compUP}
\end{align}

@@

Comparing the equivalent conditions for strong tractability in \eqref{eq:strong_tractiff3} and tractability in \eref{eq:tractiff4}, they are seen to be nearly the same.  The difference lies in the lower summation index, which is allowed to depend on $d$ for tractability, and the second argument of $T$ in the sum, which is $1$ for strong tractability, and $d$ in general.  Both of these can allow $S_{\vp}$ to be finite in Theorem \ref{thm_main_tract2}, when it may be infinite in Theorem \ref{thm_main_strong_tract2}.

@@

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,p)= \max\{1,\varepsilon^{-p}\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p \ge 0.
\]
\fred{May we pull the exponent $p$ inside the $\max$?} Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for strong algebraic polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty (1,\min\{\lambda_{i,d}^p\}) < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NW08}.
\end{example}

@@

\[
\sup_{d\in\mathbb{N}}\sum_{i=Lpd^{q_1}} \left(\min\{\lambda_{i,d},1\} \right)^{p(1+\ln d)}d^{-q} < \infty
\]
\end{example}
\kachi{
\begin{example}[Separable tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,(p,q))= \phi(\varepsilon^{-1},p)\, \psi(d,q)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p,q \ge 0.
\]
with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$. In addition, satisfying the following conditions as in (\ref{eq:ptauassume}) 
\begin{align*}
 (\phi(\varepsilon^{-1},p))^{\tau_1} &\le K_{p,\tau_1} \phi (\varepsilon^{-1},\tau_1 p)\\
  (\psi(d,q))^{\tau_2} &\le K_{q,\tau_2} \psi (d,\tau_2 q)\\
\end{align*}
where $\tau_1, \tau_2>0$. %As the name implies, the roles of $\varepsilon^{-1}$ and $d$ are separated. 
\end{example}
\begin{example}[Non-separable symmetric tractability]
    Let the tractability function, $T$, be defined by
\[
 T(\varepsilon^{-1},d,(p,q))= \exp\{\phi(\varepsilon^{-1},p)\, \psi(d,q)\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p,q \ge 0.
\]  with non-decreasing functions $f_1$, $f_2 : [1,\infty)\rightarrow[1,\infty)$.
\end{example}
}

@@

\begin{theorem}\label{thm_main_strong_tract2}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is strongly $T$-tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ and an integer $L_{\vp} > 0$ such that
\begin{equation} \label{eq:strong_tractiff3}
     S_{\vp}:=\sup_{d \in \naturals} \sum_{i = L_{\vp}}^{N(d)} \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)} < \infty.
\end{equation}
If \eqref{eq:strong_tractiff3} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tractiff3} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$.  Then $\cp_{\textup{strct}} = \widetilde{\cp}_{\textup{strct}}$, and the set of optimal strong tractability parameters is
\[
	\cp_{\textup{sopt}} =
	\{\vp^* \in \widetilde{\cp}_{\textup{strct}} :  \vp^* \notin [\widetilde{\vp},\boldsymbol{\infty}) \ \forall \widetilde{\vp}\in  \widetilde{\cp}_{\textup{strct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}

>>>>

Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},d',\vp)$ are non-decreasing in $i$ for any $d'\in\naturals$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},d',\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},d',\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},d',\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) } \le \frac{1}{T(\varepsilon^{-1},d',\vp)} \\ \forall d', n, N \in \naturals.
\end{multline}
We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer  $n - N +1$ is assumed to be positive.

Moreover, given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity via \eqref{eq:comp_def}:
\begin{align} %\label{eq:comp_equiv}
	\nonumber
	\comp(\varepsilon, d) & = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
	\nonumber
	& = \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},d',\vp)}\le \frac{1}{T(\varepsilon^{-1},d',\vp) }\biggr\} \qquad  \forall d'\in \naturals\\
	\nonumber
	& \le  \min \biggl\{n \in \natzero : n \ge N - 1 + T(\varepsilon^{-1},d',\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \biggr\} \;  \forall  d', N \in \naturals, \quad \text{by \eqref{eq:implicate_a}} \\ 
	& \le  T(\varepsilon^{-1},d',\vp) \Biggl [\frac{N}{T(\varepsilon^{-1},d',\vp)}  +  \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \Biggr] \quad \forall  d', N \in \naturals. \label{eq:compUP}
\end{align}

@@

Comparing the equivalent conditions for strong tractability in \eqref{eq:strong_tractiff3} and tractability in \eqref{eq:tractiff4}, they are seen to be nearly the same.  The difference lies in the lower summation index, which is allowed to depend on $d$ for tractability, and the second argument of $T$ in the sum, which is $1$ for strong tractability, and $d$ in general.  Both of these can allow $S_{\vp}$ to be finite in Theorem \ref{thm_main_tract2}, when it may be infinite in Theorem \ref{thm_main_strong_tract2}.

@@

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,p)= \max\{1,\varepsilon^{-p}\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p \ge 0.
\]
Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for strong algebraic polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty (1,\min\{\lambda_{i,d}^p\}) < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NW08}.
\end{example}

@@

\[
\sup_{d\in\mathbb{N}}\sum_{i=Lpd^{q_1}} \left(\min\{\lambda_{i,d},1\} \right)^{p(1+\ln d)}d^{-q} < \infty
\]
\end{example}

\kachi{In the following we introduce two specific cases of tractability which depend on the form of the function $T$.
\begin{example}[(Strong) Separable tractability]
Let the tractability function $T$,
be defined by
\[
 T(\varepsilon^{-1},d,(p,q))= \phi(\varepsilon^{-1},p)\, \psi(d,q)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p,q \ge 0.
\]
with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$. In addition, we assume the following conditions as in (\ref{eq:ptauassume}) 
%\begin{align*}
 %(\phi(\varepsilon^{-1},p))^{\tau_1} &\le K_{p,\tau_1} \phi (\varepsilon^{-1},\tau_1 p)\\ (\psi(d,q))^{\tau_2} &\le K_{q,\tau_2} \psi (d,\tau_2 q)\\\end{align*}
 \begin{align*}
     [T(\varepsilon^{-1},d,(p,q))]^\tau \leq K_{(p,q),\tau}T(\varepsilon^{-1},d,(\tau p,\tau q))
 \end{align*}
where $\tau>0$.
Observe that for this $T$, we have strong tractability if the following condition is satisfied
\[
 T(\varepsilon^{-1},1,p)= \phi(\varepsilon^{-1},p)\, \psi(1,0) = \phi(\varepsilon^{-1},p)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p \ge 0.
\]
Thus, Theorem \ref{thm_main_strong_tract2} yields the following equivalent. We say that a problem is strongly separable tractable iff there exists $p>0$ and an integer $C(p)\ge 0$ such that
\begin{equation} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L(p,1)}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\end{equation}

%As the name implies, the roles of $\varepsilon^{-1}$ and $d$ are separated. 
\end{example}
\begin{example}[Non-separable symmetric tractability]
    Let the tractability function $T$, be defined by
\[
 T(\varepsilon^{-1},d,(p,q))= \exp\{\phi(\varepsilon^{-1},p)\, \psi(d,q)\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p\ge 0.
\]  with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$.
\end{example}
}

@@

\begin{theorem}\label{thm_main_strong_tract3}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is strongly $T$-tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ and an integer $L_{\vp} > 0$ such that
\begin{equation} \label{eq:strong_tractiff3}
     S_{\vp}:=\sup_{d \in \naturals} \sum_{i = L_{\vp}}^{N(d)} \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)} < \infty.
\end{equation}
If \eqref{eq:strong_tractiff3} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tractiff3} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$.  Then $\cp_{\textup{strct}} = \widetilde{\cp}_{\textup{strct}}$, and the set of optimal strong tractability parameters is
\[
	\cp_{\textup{sopt}} =
	\{\vp^* \in \widetilde{\cp}_{\textup{strct}} :  \vp^* \notin [\widetilde{\vp},\boldsymbol{\infty}) \ \forall \widetilde{\vp}\in  \widetilde{\cp}_{\textup{strct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}
